---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  include = TRUE,
  warning = FALSE,
  fig.width = 6,
  fig.height = 5,
  fig.align = 'center',
  comment = "#>",
  fig.path = "man/figures/README-"
  )
library(fwa.connect)
```

# fwa.connect

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/fwa.connect)](https://CRAN.R-project.org/package=fwa.connect)
<!-- badges: end -->

Fwa.connect is intended to help users quickly identify patterns of connectivity between streams and lakes in the B.C. Data Catalogue's Freshwater Atlas (henceforth, 'FWA'). By providing a two-column data.table that lists up- and down-stream streams (identified by the FWA_WATERSHED_CODE field), a tidygraph network object, and some basic utility functions to work with upstream graphs of streams, {fwa.connect} will hopefully reduce workflow waiting times by eliminating the need to download the entirety of the FWA stream network or to perform laborious spatial operations to find connections between streams.

## Installation

You can install the development version of fwa.connect like so:

``` r
devtools::install_github('chrispmad/fwa.connect')
# or 
remotes::install_github('chrispmad/fwa.connect')
```

## Functions

### Included dataset: fwa_up_and_downstream_tbl

A two-column table from which an {igraph} / {tidygraph} graph object can be derived. The first column represents the FWA_WATERSHED_CODE (near-unique ID column) of a target stream in the FWA ('upstream_fwa_code'), the second represents the FWA_WATERSHED_CODE of the stream downstream of the target stream.
```{r raw_data_table}
knitr::kable(head(fwa.connect::fwa_up_and_downstream_tbl, 2))
```

### fwa_graph

This function takes the included dataset described above and turns it into a {tidygraph} graph object.
```{r fwa_graph}
if(FALSE) fwa.connect::fwa_graph() # tbl_graph with 1523261 nodes and 1522833 edges. 
# Not sure how to display this in brief, so it's quarantined behind an 'if(FALSE)' statement for our safety!
```

### trace_course_downstream

Trace the course of flow downstream from the stream you identify with its FWA WATERSHED CODE id. This function returns an {sf} spatial table and an optional ggplot.
```{r trace_course_downstream_eg}
# An example FWA code.
fwa_code = "200-948755-999851-274772-093336-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000"

ds = fwa.connect::trace_course_downstream(fwa_code = fwa_code,
                        merge_by_BLK = T,
                        make_plot = T, 
                        add_map_insert = T,
                        save_plot = F,
                        save_plot_location = NA)

knitr::kable(head(ds$downstream_course, 2))

ds$plot
```

### trace_course_upstream

```{r trace_course_upstream_eg}
# An example FWA code.
fwa_code = "200-948755-999851-274772-093336-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000-000000"

us = fwa.connect::trace_course_upstream(fwa_code = fwa_code,
                        merge_by_BLK = T,
                        make_plot = T, 
                        add_map_insert = T,
                        save_plot = F,
                        save_plot_location = NA)

knitr::kable(head(us$upstream_streams, 2))

us$plot
```

### estimate_total_upstream_length 

Estimate the summed lengths of all streams above a certain point in space, e.g. a barrier to fish passage. This can be done for a single point or for an {sf} table of multiple points.

#### Single point

```{r est_length_single_point_example}
# library(bcdata) # To access datasets in the open-source BC Data Catalogue
# library(sf)
# library(progress)

# Download a possible fish barrier from the PSCIS dataset.

# This is the code for pscis-assessments:
pscis_layer_id = '7ecfafa6-5e18-48cd-8d9b-eae5b5ea2881'

fp = bcdata::bcdc_query_geodata(pscis_layer_id) |> 
  bcdata::filter(RESPONSIBLE_PARTY_NAME == 'WEST FRASER MILLS LTD.',
         STREAM_NAME == 'Nass River',
         ROAD_NAME == 'Warren Road') |>
  bcdata::collect()

# Find the nearest stream within 50 meters.
stream = fwa.connect::find_nearest_stream(fp, max_buffer_dist = 50)

# Calculate the summed length of all streams upstream from a point (or a FWA code)
upstream_l = fwa.connect::estimate_total_upstream_length(obstacles = fp,
                                            make_plot = T,
                                            save_plot = F)

knitr::kable(
  upstream_l 
)
```

#### Multiple points

```{r est_length_multiple_points_example}
# Calculate the length for multiple points.
fps = bcdata::bcdc_query_geodata(pscis_layer_id) |>
  bcdata::filter(ASSESSMENT_DATE > as.Date('2020-10-01') & ASSESSMENT_DATE < as.Date('2021-01-01')) |> 
  bcdata::collect()

upstream_lengths = fwa.connect::estimate_total_upstream_length(
  obstacles = fps,
  make_plot = F,
  save_plot = F
)

knitr::kable(
  upstream_lengths
)
```
