---
title: "my_vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my_vignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = F,
  warning = F,
  fig.width = 8,
  fig.height = 5
)
library(fwa.connect)
```

## Foreward

This package, {fwa.connect}, is intended to make your life easier when working with the British Columbia Freshwater Atlas (FWA); specifically, the stream network line layer and the lakes layer. These layers (and perhaps others in the FWA) share a column: the FWA_WATERSHED_CODE. This column is quite ingenious, as it succinctly performs at least four functions:
1. It serves as an (almost) unique ID column.
2. For a given stream X, it describes how far down the 'receiving' stream Y the stream X joins onto stream Y.
3. It implies how many streams or rivers away from the massive watershed scale collecting river a target stream is: e.g., a FWA code of 100-274173-283712-859301-000000... implies that our stream makes up a chain of four connected streams, including the massive collecting river (the '100', in this case).
4. It can be used to generate a directed 'graph', which we can then use to quickly calculate useful graph attributes for each node (i.e. stream) in the graph, e.g. centrality, component membership (groups of nodes connected by at least one edge), betweenness, shortest path between two points, eccentricity, etc.

## Grab Toy Dataset

```{r get_toy_dataset}
# ---------------
# Get test points

# Prince George Natural Resource District Polygon
dpg = bcmaps::nr_districts() |>
  dplyr::filter(ORG_UNIT_NAME == 'Prince George Natural Resource District')

# Potential barriers to fish movement in streams
fps_DPG = bcdata::bcdc_query_geodata("pscis-assessments") |>
  bcdata::filter(INTERSECTS(dpg)) |>
  bcdata::filter(ASSESSMENT_DATE > as.Date('2023-09-01') & ASSESSMENT_DATE < as.Date('2024-12-31')) |>
  #bcdata::filter(CURRENT_BARRIER_RESULT_CODE = "BARRIER") |>
  bcdata::collect()

# Single barrier point
fps_single = fps_DPG |>
  dplyr::filter(STREAM_CROSSING_ID == 198771)

# Multiple (N = 100) barrier points
fps_multi = sample(fps_DPG, size = 20, replace = F)
```

## Functions

The principle functions of the {fwa.connect} package are as follows:

### 1. Freshwater Atlas Stream Network connectivity (*fwa_graph*)

- packages the FWA stream network as a two-column table describing source and destination rivers/streams (N = 1,523,261). This can be easily converted into a {tidygraph} graph object, with a table of 1.5 million nodes and a table of edge data (connections between nodes; N = 1,522,833). This graph supports many subsequent functions and can be used by any savvy analyst that knows the basics of {igraph} and/or {tidygraph}.

```{r}
str_tbl = stream_conn_tbl()

head(str_tbl)

tidygraph::as_tbl_graph(str_tbl)
```

### 2. Delineate Component Groups (*delineate_comps*)

- Identifies the component membership of one or more FWA_WATERSHED_CODEs.

```{r delineate_comps_example}
# Just to ensure we have streams that are part of different groups, let's snag the geometries of three different streams around the province.
stream_network_id = '92344413-8035-4c08-b996-65a9b3f62fca'

some_streams = bcdata::bcdc_query_geodata(stream_network_id) |> 
  bcdata::filter(GNIS_NAME %in% c("Paul River","Spius Creek","Taku River")) |> 
  bcdata::collect()

some_streams = delineate_comps(some_streams)

library(ggplot2)

ggplot() + 
  geom_sf(data = bcmaps::bc_bound()) +
  geom_sf(data = some_streams, aes(col = comp_group, fill = comp_group)) +
  labs(title = 'These three streams are part of separate graph components')
```

### 3. Find Nearest Stream (*find_nearest_stream*)

- Finds the nearest stream for an {sf} POINT simple feature collection of 1 or more features.

```{r nearest_stream_single_point_example}
# Add FWA_WATERSHED_CODE to table of points and also return stream geometries
fps_single_l = find_nearest_stream(fps_single)

# Unpack fps_DPG and streams
fps_single_p = fps_single_l$points
fps_single_s = fps_single_l$streams

# How many barriers did we find a sufficiently close stream for?
nrow(fps_single_p |>
  dplyr::filter(!is.na(fwa_code)))

# Geometry of nearest stream to obstacle.
plot(fps_single_s$geometry)
plot(fps_single_p$geometry, add = T, col = 'red', lwd = 3)

```

```{r nearest_stream_multi_point_example}
# Add FWA_WATERSHED_CODE to table of points and also return stream geometries
fps_multi_l = find_nearest_stream(fps_multi)

# Unpack fps_DPG and streams
fps_multi_p = fps_multi_l$points
fps_multi_s = fps_multi_l$streams

# How many barriers did we find a sufficiently close stream for?
nrow(fps_multi_p |>
  dplyr::filter(!is.na(fwa_code)))

# Geometry of nearest stream to obstacle.
plot(fps_multi_s$geometry)
plot(fps_multi_p$geometry, add = T, col = 'red', lwd = 3)
```

### 3. Trace Course Downstream (*trace_course_downstream*)

- Finds collecting streams downstream of one or more points or streams.

```{r}
minaker = bcdata::bcdc_query_geodata('freshwater-atlas-stream-network') |> 
  bcdata::filter(GNIS_NAME == 'Minaker River') |> 
  bcdata::collect() |> 
  sf::st_zm() |> 
  dplyr::group_by(GNIS_NAME, BLUE_LINE_KEY, FWA_WATERSHED_CODE) |> 
  dplyr::summarise()

trace_course_downstream(
  minaker$FWA_WATERSHED_CODE,
  make_plot = T,
  add_map_insert = T
)

# Not working for multiple points yet.
# trace_course_downstream(
#   fps_multi_p$fwa_code,
#   make_plot = T
# )
```

### 4. Trace Course Upstream (*trace_course_upstream*)

- Finds all streams upstream of one or more streams. Optional ggplot that shows the submitted streams in dark blue. Optional merging of streams in output table by BLUE_LINE_KEY and a handful of other columns.

```{r}
# Single point
fps_s_upstr = trace_course_upstream(
  fps_single_s$FWA_WATERSHED_CODE,
  make_plot = T,
  add_map_insert = T
)

fps_s_upstr$plot

# Multiple points
fps_m_upstr = trace_course_upstream(
  fps_multi_s$FWA_WATERSHED_CODE,
  make_plot = T,
  add_map_insert = T
)

fps_m_upstr$plot
```


### 5. Estimate Upstream Stream Length (*estimate_total_upstream_length()*)

Use the estimate function for a single barrier:

```{r, fig.width=6, fig.height=6}
obst_with_length = fwa.connect::estimate_total_upstream_length(
  fps_DPG[1,],
  make_plot = T,
  stream_snap_dist = 80,
  min_obstacles_separation = 40
)

obst_with_length$total_length_m
```

Or use the estimate function for multiple barriers in one function call

```{r}
multi_obst_with_l = estimate_total_upstream_length(
  fps_DPG[c(1:10),],
  make_plot = F,
  stream_snap_dist = 80,
  min_obstacles_separation = 40
)

multi_obst_with_l$total_length_m
```
